---
title: "Lab"
author: "Namuun"
date: "2023-12-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
library(magrittr)
install.packages("dplyr")
library(dplyr)
install.packages("ggplot2") 
library(ggplot2)

## R Assignment 1

* In this assignment, we are going to apply some of the build in data set in R for descriptive statistics analysis.
* To earn full grade in this assignment, students need to complete the coding tasks for each question to get the result.
* After finished all the questions, knit the document into HTML format for submission.

### Question 1

Using the **mtcars** data set in R, please answer the following questions.

```{r}
# Loading the data
data(mtcars)

# Head of the data set
head(mtcars)
```

a. Report the number of variables and observations in the data set.

```{r}
# Enter your code here!
num_variables <- ncol(mtcars)
num_observations <- nrow(mtcars)

# Answer:
print(paste("There are total of", num_variables, "variables and", num_observations, "observations in this data set."))
```



b. Print the summary statistics of the data set and report how many discrete and continuous variables are in the data set.

```{r}
# Enter your code here!
summary_stats <- summary(mtcars)
num_discrete <- sum(sapply(mtcars, is.factor))
num_continuous <- num_variables - num_discrete

# Answer:
print(paste("There are", num_discrete, "discrete variables and", num_continuous, "continuous variables in this data set."))
```



c. Calculate the mean, variance, and standard deviation for the variable **mpg** and assign them into variable names m, v, and s. Report the results in the print statement.

```{r}
# Enter your code here!
mpg_mean <- mean(mtcars$mpg)
mpg_var <- var(mtcars$mpg)
mpg_sd <- sd(mtcars$mpg)



# print(paste("The average of Mile Per Gallon from this data set is ",  , " with variance ",  , " and standard deviation",  , "."))
print(paste("The average of Mile Per Gallon from this data set is", mpg_mean, "with variance", mpg_var, "and standard deviation", mpg_sd, "."))

```



d. Create two tables to summarize 1) average mpg for each cylinder class and 2) the standard deviation of mpg for each gear class.

```{r}
# Enter your code here!
mean_mpg_by_cyl <- aggregate(mpg ~ cyl, mtcars, mean)
sd_mpg_by_gear <- aggregate(mpg ~ gear, mtcars, sd)

print(mean_mpg_by_cyl)
print(sd_mpg_by_gear)


```



e. Create a crosstab that shows the number of observations belong to each cylinder and gear class combinations. The table should show how many observations given the car has 4 cylinders with 3 gears, 4 cylinders with 4 gears, etc. Report which combination is recorded in this data set and how many observations for this type of car.

```{r}
# Enter your code here!
cyl_gear_crosstable <- table(mtcars$cyl, mtcars$gear)
print(cyl_gear_crosstable)

print("The most common car type in this data set is car with 8 cylinders and 3 gears. There are total of 12 cars belong to this specification in the data set.")
```


*** 

### Question 2

Use different visualization tools to summarize the data sets in this question.

a. Using the **PlantGrowth** data set, visualize and compare the weight of the plant in the three separated group. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your findings.

```{r}
# Load the data set
data("PlantGrowth")

# Head of the data set
head(PlantGrowth)

# Enter your code here!
# Enter your code here!

library(ggplot2)
ggplot(PlantGrowth, aes(x = group, y = weight)) + 
    geom_boxplot() +
    labs(title = "Comparison of Plant Weights",
         x = "Group",
         y = "Weight") +
    theme_minimal()

```

Result:

=> Report a paragraph to summarize your findings from the plot!
The median weight for the control group is slightly above 5.0, and there are no outliers. Treatment 1 also has a comparable median weight, but it includes two notable outliers, suggesting that certain plants experienced substantially higher weights. In Treatment 2, the median weight is higher, around 5.3, and there is a broader range, indicating more consistent increases in plant weight. In summary, Treatment 2 seems to have the most beneficial effect on plant weight.

b. Using the **mtcars** data set, plot the histogram for the column **mpg** with 10 breaks. Give labels to the title, x-axis, and y-axis on the graph. Report the most observed mpg class from the data set.

```{r}
hist(mtcars$mpg, breaks = 10, main = "Histogram of MPG", xlab = "Miles Per Gallon (MPG)", ylab = "Frequency")


print("Most of the cars in this data set are in the class of 15 miles per gallon.")

```



c. Using the **USArrests** data set, create a pairs plot to display the correlations between the variables in the data set. Plot the scatter plot with **Murder** and **Assault**. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your results from both plots.

```{r}
# Load the data set
data("USArrests")

# Head of the data set
head(USArrests)

# Enter your code here!
pairs(USArrests, main = "Pairs Plot of USArrests Data")
plot(USArrests$Murder, USArrests$Assault, main = "Scatter Plot of Murder vs Assault",
     xlab = "Murder Rate", ylab = "Assault Rate")
```

Result:

=> Report a paragraph to summarize your findings from the plot!

The scatter plot indicates a positive connection between Murder Rate and Assault Rate—when one goes up, the other tends to rise as well. The majority of points cluster in the lower left, showing that most states have lower rates for both murders and assaults. While a few states have higher rates for both, such instances are infrequent. The plot doesn't reveal any outliers with exceptionally high or low rates compared to the general trend.

***

### Question 3

Download the housing data set from www.jaredlander.com and find out what explains the housing prices in New York City.

Note: Check your working directory to make sure that you can download the data into the data folder. 

```{r, echo=FALSE}
# Load and clean the housing data set 
download.file(url='https://www.jaredlander.com/data/housing.csv',
              destfile='data/housing.csv', mode='wb')
housingData <- read.csv('data/housing.csv')
housingData <- subset(housingData, 
                      select = c("Neighborhood", "Market.Value.per.SqFt", "Boro", "Year.Built"))
housingData <- na.omit(housingData)                                     
```


a. Create your own descriptive statistics and aggregation tables to summarize the data set and find any meaningful results between different variables in the data set.

```{r}
# Head of the cleaned data set
head(housingData)

# Enter your code here!
summary(housingData)
agg1 <- aggregate(Market.Value.per.SqFt ~ Year.Built, data = housingData, mean)

```

b. Create multiple plots to demonstrates the correlations between different variables. Remember to label all axes and give title to each graph.

```{r}
# Enter your code here!
plot(agg1$Year.Built, agg1$Market.Value.per.SqFt, 
     main = "Average Market Value per Square Foot by Year",
     xlab = "Year Built", ylab = "Average Market Value per Square Foot", pch = 19)
boxplot(housingData$Market.Value.per.SqFt ~ housingData$Boro, 
        main = "Box Plot of Market Value per SqFt by Borough",
        xlab = "Borough", 
        ylab = "Market Value per SqFt")

```

c. Write a summary about your findings from this exercise.

=> Enter your answer here!

Scatter Plot:
The scatter plot suggests that the market value per square foot varies widely for properties built between 1850 and 2000. Construction year doesn't strongly predict value.

Box Plots:
Manhattan has the highest median property value, followed by Brooklyn, Queens, and the Bronx. Manhattan has the most diverse property values, while the Bronx has the lowest median. Outliers in all boroughs indicate some properties significantly differ from the median.